---
title: "Introduction to shiny.i18n"
author: "Appsilon"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to shiny.i18n}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, screenshot.force = FALSE}
library(shiny)
library(shiny.i18n)
```


In this tutorial we will walk you through the basics of `shiny.i18n` (which means internationalization of `Shiny`).
The package can be used in 3 modes.
Although it was designed to work smoothly with `Shiny` apps and reactivity, it can be also
used for translation of other R documents, for example RMarkdown. Let's have a look at the 3 most
common use-cases of translation.

## Back-end translation

Imagine that you have a Shiny app like the one below.

```{r echo = T, include = T, eval = FALSE}
library(shiny)
library(shiny.i18n)

ui <- fluidPage(p("Hello"))

server <- function(input, output) {}

shinyApp(ui, server)
```

Now, you would like to translate the content of your application using the translations
from a file *translation.json*.

**More about the format of translation file you can find in the other tutorial**.

First of all, you need to load the data and create Translator object:

```{r echo = T, include = T, eval = FALSE}
i18n <- Translator$new(translation_json_path = "translation.json")
```

Next just surround each text in your app by `i18n$translate` or `i18n$t` (for short) tags. For example:

```{r echo = T, include = T, eval = FALSE}
i18n$translate("Hello")
# or
i18n$t("Hello")
```

The only thing that is left is selecting the target translation language. You do it with
`i18n$set_translation_language` method that takes as an argument the code of your language.

Your full application might look now like this:

```{r echo = T, include = T, eval = FALSE}
library(shiny)
library(shiny.i18n)

i18n <- Translator$new(translation_json_path = "translation.json")

i18n$set_translation_language("en")

ui <- fluidPage(p(i18n$t("Hello")))

server <- function(input, output) {}

shinyApp(ui, server)
```

If you want to display the text in a different language, just change the code from "en"
to the language that is in your *"translation.json"* file, eg. "pl" for Polish.

## Live language change

The example above demonstrates only how to run various instances of the app
with different languages. What if we want to enable user to change the language dynamically?

You have two options here.

### Option (a)

You can make your i18n a reactive variable that accepts the language code from
some UI input, eg. `input$selected_language`.

```{r echo = T, include = T, eval = FALSE}
translator <- Translator$new(translation_json_path = "translation.json")

server <- function(input, output, session) {

  i18n <- reactive({
    selected <- input$selected_language
    if (length(selected) > 0 && selected %in% translator$get_languages()) {
      translator$set_translation_language(selected)
    }
    translator
  })
}
```

See the full example [here](https://github.com/Appsilon/shiny.i18n/blob/master/examples/live_language_change/server_app.R).

### Option (b)

Alternatively, you may introduce to your UI the following statement: `usei18n(i18n)`. This will tell
`shiny.i18n` to use JS script nested in the header of your app to perform the translation.

See this minimal example:

```{r echo = T, include = T, eval = FALSE}
library(shiny)
library(shiny.i18n)

i18n <- Translator$new(translation_json_path = "translation.json")

i18n$set_translation_language("en")

ui <- fluidPage(usei18n(i18n),
                p(i18n$t("Hello")),
                actionButton("go", "GO!")
      )

server <- function(input, output, session) {
  observeEvent(input$go,{
       update_lang(session, "pl")
  })
}

shinyApp(ui, server)
```

## Automatic translations

Sometimes there is no time to translate your application to many different languages
and there is a number of automatic translation services with public API. With `shiny.i18n`
it's easy to take advantage of some of them. Remember though that you need to set-up
credential to the service you want to use first. Here is an example on how to
internationalize your app with Google API (note that we replace `i18n$t` by `i18n$at`).

**(! Note that you need to check with a provider of the API for a specific language code)**

```{r echo = T, include = T, eval = FALSE}
library(shiny)
library(shiny.i18n)
library(googleLanguageR)


# setting up credentials to google cloud API (see more in googleLanguageR docs)
i18n <- Translator$new(automatic = TRUE)

i18n$set_translation_language("de")

ui <- fluidPage(p(i18n$at("Hello")))

server <- function(input, output) {}

shinyApp(ui, server)
```

That's the end of this tutorial. Hopefully you learned something new about `shiny.i18n`.
Check more examples in `examples` folder of `shiny.i18n` code [repository](https://github.com/Appsilon/shiny.i18n).
